---
title: "Tidy Tuesday"
subtitle: "National Parks Species Data"
author: "Hanna Chang"
format: 
  html:
    embed-resources: true
---

## Loading packages

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(sf)
library(scales)
```

## Loading data

```{r, message=FALSE, warning=FALSE}
most_visited_nps_species_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-10-08/most_visited_nps_species_data.csv')
```

# With a glance at the initial data, which is huge, I'd like to explore the different species that each national park has. Maybe I can also see how diverse the national parks are.

## Data wrangling by name of national park

```{r, message=FALSE, warning=FALSE}
by_parkname <- most_visited_nps_species_data %>%
  group_by(ParkName) %>%
  summarise(total_species = n_distinct(SciName)) %>%
  arrange(desc(total_species))

head(by_parkname, 10)
```

## More data wrangling for each national park's species, grouped by category name

```{r, message=FALSE, warning=FALSE}
by_category <- most_visited_nps_species_data %>%
  group_by(CategoryName) %>%
  summarise(total_species = n_distinct(SciName)) %>%
  arrange(desc(total_species))

join_category <- most_visited_nps_species_data %>%
  left_join(by_category, by = "CategoryName") %>%
  group_by(ParkName, CategoryName) %>%
  summarise(total_species = n_distinct(SciName)) %>%
  arrange(desc(total_species))

join_category <- join_category %>%
  group_by(ParkName) %>%
  mutate(percentage = total_species / sum(total_species) * 100)
```

# Each national park's species-diversity pie chart

```{r, fig.width=15, fig.height=10, message=FALSE, warning=FALSE}
ggplot(join_category, aes(x = "", y = percentage, fill = CategoryName)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  facet_wrap(~ParkName, ncol = 5) +
  theme_void() +
  theme(legend.position = "bottom", legend.text = element_text(size = 10),
        legend.title = element_text(size = 12), legend.key.size = unit(0.8, "cm"),
        strip.text = element_text(size = 9.5),
        plot.title = element_text(hjust = 0.5), panel.spacing = unit(1.5, "lines"),
        margin = margin(b = 20)) +
  scale_fill_viridis_d(option = "G") +
  labs(title = "Each National Park's species diversity by category name")
```

## Based on the visual, I'd like to see if species diversity has anything to do with the park's visitor count. I could investigate that in the future.








